Searching for Software in NSF Awards

### 2023-02-09 paper/paper.qmd
Tiny bit more work on paper (models section)

from soft_search.data.irr import print_irr_summary_stats
from IPython.display import Markdown

@@

from soft_search.data import load_soft_search_2022
from soft_search.data.soft_search_2022 import SoftSearch2022DatasetFields
from soft_search.constants import PredictionLabels
from IPython.display import Markdown

@@

We trained multiple predictive models using different model architectures and using the two different predictive sources available (the award's abstract text or project outcomes report text). As not all awards in our set had concluded their research, the models trained on the project outcomes report text used a reduced size dataset.

We trained three different models for each of the predictive sources and include a custom regular expression as a baseline for comparison. The models trained include a TF-IDF Word Embedding Logistic Regression model, a Semantic Embedding Logistic Regression model, and a Fine-tuned Transformer model.

@@

model_results = fit_and_eval_all_models(train_transformer=False)
```

```{python}
#| code-fold: true
#| warning: false
#| label: tbl-models-from-abstract
#| tbl-cap: Model Results from Training using Abstract Text

model_results[model_results["predictive_source"] == "abstract-text"].reset_index(drop=True)
```

```{python}
#| code-fold: true
#| warning: false
#| label: tbl-models-from-project-outcomes
#| tbl-cap: Model Results from Training using Project Outcomes Reports Text

model_results[model_results["predictive_source"] == "project-outcomes"].reset_index(drop=True)
```

>>>>

from soft_search.data.irr import print_irr_summary_stats

@@

from soft_search.data import load_soft_search_2022
from soft_search.data.soft_search_2022 import SoftSearch2022DatasetFields
from soft_search.constants import PredictionLabels

@@

We trained three different models using the text from either the award abstract or project outcomes report. The models trained include a logistic regression model trained with TF-IDF word embeddings (`tfidf-logit`), a logistic regression model trained with semantic embeddings (`semantic-logit`), and a fine-tuned transformer (`fine-tuned-transformer`). The semantic embeddings and the base model from which we fine-tuned our own transformer model was the `distilbert-base-uncased-finetuned-sst-2-english` model [@hf_canonical_model_maintainers_2022]. Additionally, we evaluated a regular expression which matched on common terms relating to software as a baseline.

@@

model_results = fit_and_eval_all_models()
```

```{python}
#| code-fold: true
#| warning: false
#| label: tbl-model-results-from-abstract
#| tbl-cap: Predictive Model Results (Trained with Abstract Text)

from_abstract = model_results[
  model_results["predictive_source"] == "abstract-text"
].reset_index(drop=True)

from_abstract
```

```{python}

best_model_from_abstract = from_abstract.iloc[0]
best_model_from_abstract_name = best_model_from_abstract["model"]
best_model_from_abstract_f1 = best_model_from_abstract["f1"]

Markdown(
  f"@tbl-model-results-from-abstract reports the results from training using "
  f"the abstract text as input. The best performing model was the "
  f"`{best_model_from_abstract_name}` which achieved an F1 of "
  f"{round(best_model_from_abstract_f1, 3)}."
)
```

```{python}
#| code-fold: true
#| warning: false
#| label: tbl-model-results-from-project-outcomes
#| tbl-cap: Predictive Model Results (Trained with Project Outcomes Report Text)

from_po = model_results[
  model_results["predictive_source"] == "project-outcomes"
].reset_index(drop=True)

from_po
```

```{python}

best_model_from_po = from_po.iloc[0]
best_model_from_po_name = best_model_from_po["model"]
best_model_from_po_f1 = best_model_from_po["f1"]

Markdown(
  f"@tbl-model-results-from-project-outcomes reports the results from training using "
  f"the project outcomes reports as input. The best performing model was the "
  f"`{best_model_from_po_name}` which achieved an F1 of "
  f"{round(best_model_from_po_f1, 3)}."
)
```
